---
title: "HW_1.5"
author: "Group name: Chris and Earl's Econometric Adventure"
date: "Sep 6, 2018"
output:
  pdf_document: default
---
\newpage

```{r}
library(scales)
```

## Appendix A
**Questions: 2,4,7 (all parts), 10 (parts i and iii)**
\newline

2. Suppose the following equation describes the relationship between the average number of classes missed during a semester(*missed*) and the distance from school (*distance*, measured in miles): $$ \textit{missed} = 3 + 0.2 \textit{distance} $$
\newline

(i) Sketch this line, being sure to label the axes. How do you interpret the intercept in this equation?
\newline
```{r}
distance <- c(0:30)
missed <- 3 + 0.2*distance
plot(distance,missed, xlim = c(1,30), ylim = c(1,30), xlab = "Distance From School", ylab = "Days Missed")
abline(3,0.2, col = "red")
```

**Answer\:** The intercept in this equation is 3. This implies that a student is likely to miss at least 3 days regardless of the distance they live from school
\newline

(ii) What is the average number of classes missed for someone who lives five miles away?
\newline
```{r}
fivemilesaway <- 3 + 0.2*5

```

**Answer\:** The average number of classes missed for someone who lives five miles away is `r fivemilesaway` classes.
\newline

(iii) What is the difference in the average number of classes missed for someone who lives 10 miles away and someone who lives 20 miles away?
\newline
```{r}
tenmilesaway <- 3 + 0.2 * 10
twentymilesaway <- 3 + 0.2 * 20
difference <- twentymilesaway - tenmilesaway
```


**Answer\:** The difference in the average number of classes missed for someone who lives 10 miles away and someone who lives 20 miles away is `r difference` classes.
\newline

4. Suppose the unemployment rate in the United States goes from 6.4% in one year to 5.6% in the next. 
\newline

(i) What is the percentage point decrease in the unemployment rate?
\newline

**Answer\:** The unemployment rate in the United States decreased by `r 6.4 - 5.6`%.
\newline

(ii) By what percentage has the unemployment rate fallen?
\newline

**Answer\:** The percentage decrease in the unemployment rate is `r -1 * ((5.6 - 6.4)/6.4)*100`%.
\newline

7. Suppose the following model describes the relationship between annual salary (*salary*) and the number of previous years of labor market experience (*exper*): $$ log(\textit{salary}) = 10.6 + 0.027 \textit{exper} $$
\newline

$$ \textit{salary} = e{10.6 +0.027\textit{exper}}$$
```{r}

exper <- c(0:30)
exponent <- 10.6 + 0.027*exper
salary <- exp(exponent)

exper
salary
plot(exper,salary, xlim = c(1, 30))
```


(i) What is *salary* when *exper* = 0? When *exper* = 5? (*Hint: You will need to exponentiate.*)
\newline

```{r}
exper0 <- 0
salary0 <- exponent <- exp(10.6 + 0.027*exper0)

exper5 <- 5
salary5 <- exponent <- exp(10.6 + 0.027*exper5)
```


**Answer\:** The salary when exper = 0 is `r dollar_format()(salary0)`. The salary when exper = 5 is `r dollar_format()(salary5)`
\newline

(ii) Use equation (A.28) to approximate the percentage increase in *salary* when *exper* increases by 5 years.
\newline
$$ \% \Delta y \approx (100 * \beta_1) \Delta x$$
```{r}
deltay <- (100 * 0.027) * 5
```

**Answer\:** The approximate increase in *salary* when *exper* increases by 5 years is `r deltay`%.
\newline

(iii) Use the results of part (i) to compute the exact percentage difference in salary when *exper* = 5 and *exper* = 0. Comment on how this compares with the approximation in part (ii).
\newline

```{r}
deltaactual <- 100* ((salary5 - salary0)/salary0)
```

**Answer\:** The exact percentage difference in salary when *exper* = 5 and *exper* = 0 is `r deltaactual`%. This is a difference of `r -1*(deltay - deltaactual)`%.
\newline

10. Suppose that in a particular state a standardized test is given to all graduating seniors. Let *score* denote a student's score on the test. Someone discovers that performance on the test is related to the size of the student's graduating high school class. The relationship is quadratic: $$ \textit{score} = 45.6 + 0.82\textit{class} - 0.000147\textit{class}^2 $$

where *class* is the number of students in the graduating class. 
\newline

```{r}
classsize <- c(0:5000)
score <- 45.6 + 0.82*classsize - (0.000147)*(classsize^2)
plot(classsize, score, xlim = c(1,5000), ylim = c(1, 3000))

score_derivative <- .82/(0.000147*2)
maxscore <- max(score)
maxscore
score_derivative
```


(i) How do you literally interpret the value 45.6 in the equation? By itself, is it of much interest?
\newline

**Answer\:** 
\newline

(iii) Sketch a graph that illustrates your solution in part (ii).
\newline

**Answer\:**
\newline


## Appendix B
**Questions: 2,4,9, and 10**
\newline

2. Let X be a random variable distributed as Normal(5,4). Find the probabilities of the following events: 
\newline

```{r}
mean2 <- 5
sd2 <- 4
```


(i) $P(X \leq 6)$
```{r}
zi <- (6-5)/4
```


**Answer\:** The probability of X $\leq$ 6 gives a z-score of `r zi` this gives the probability of 59.87%.
\newline

(ii) $P(X > 4)$
\newline

```{r}
zii <- (4-5)/4
```


**Answer\:** The probability of X > 4 gives a z-score of `r zii` this gives the probability of 59.87%.
\newline

(iii) $P(\mid X - 5 \mid > 1)$
\newline
This means we want: $$P(X - 5 > 1) + P(X - 5 < -1)$$ $$P(X > 6) + P(X < 4)$$
```{r}
ziii6 <- (6 - 5)/4
ziii4 <- (4 - 5)/4

ziii6table <- .5987
ziii4table <- .4013

iiiprob <- (1 - ziii6table) + ziii4table
```


**Answer\:** This probability for:$P(X > 6) + P(X < 4)$ is `r iiiprob*100`%.
\newline

4. For a randomly selected county in the United States, let X represent the proportion of adults over age 65 who are employed, or the elderly employment rate. Then, X is restricted to a value between zero and one. Suppose that the cumulative distribution function for X is given by: $$F(x) = 3x^2 - 2x^3 \;\;\;\;\;\;\;\;\;\; for\;\;\; 0 \leq x \leq 1$$
Find the probability that the elderly employment rate is at least .6 (60%).
\newline

```{r}
elderemp <- seq(0,1,.01)
distri <- 3*(elderemp**2) - 2*(elderemp**3)
plot(elderemp,distri)
```


**Answer\:**
\newline

9. Let X denote the annual salary of university professors in the United States, measured in thousands of dollars. Suppose that the average salary is 52.3, with a standard deviation of 14.6. Find the mean and standard deviation when salary is measured in dollars.
\newline

**Answer\:**
\newline

10. Suppose that at a large university, college grade point average, *GPA*, and SAT score *SAT*, are related by the conditional expectation: $$E(\textit{GPA}|\textit{SAT}) = 0.70 + 0.002\textit{SAT}$$
\newline

(i) Find the expected *GPA* when *SAT* = 800. Find: $$(\textit{GPA}|\textit{SAT}) = 1400$$. Comment on the difference.
\newline

**Answer\:**
\newline

(ii) If the average *SAT* in the university is 1,100, what is the average *GPA* (*Hint:* Use Property CE.4)
\newline

**Answer\:**
\newline

(iii) If a student's SAT score is 1,100, does this mean he or she will have the GPA found in part (ii)? Explain.
\newline

**Answer\:**
\newline



